---
title: "GG606: Analytics Demo"
author: "Julia Kozak"
date: "April 04, 2024"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float:
      collapsed: false
      smooth_scroll: false
    highlight: kate
    number_sections: false
    keep_md: true
  pdf_document:
    toc: true
    toc_depth: '6'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**This assignment follows two roads within each section: the easy way and the hard way. This is because I am not sure what the authors did exactly to receive these results. The 'easy' way follows supplementary data provided outside of the data frame download, which retrieves the results the authors received post LME modeling. The hard way is explicitly taking the raw data `.csv` file and attempting to recreate their LME model. However, because portions of crucial information is missing from the process, guesstimate work was conducted in an attempt to achieve the same final results.**

________________________________________________________________________________
## 1. Loading Things: \

#### 1.1: Loading Packages \
Load in any and all packages required to manipulate, transform, and illustrate the data given.
```{r Loading Packages, include=TRUE}
#Functional:
library(tidyverse)     #Package to process our data, stylized formation
library(magrittr)      #Package to help coding sequencing 
library(janitor)       #Package for 'clean_names()' function
library(Matrix)        #Package to run `lme4` package functions
library(nlme)          #Package for LME modeling, specific for `predict_se()`
library(lme4)          #Package to plot mixed linear and non-linear models
library(emmeans)       #Package for least squares means, needed for LME stats
library(AICcmodavg)    #Package to determine SE error values
#Aesthetics:
library(ggplot2)       #Package to generate plotted data
library(patchwork)     #Package for extensive plotted data configuration 
library(ggthemes)      #Package for extra themes, scales, and geoms for plotted data
library(RColorBrewer)  #Package to colour plots

library(here)          #Package to set working directory via `.Rproj`
getwd()                #Function to affirm working directory 
```

#### 1.2: Loading Data \
The only data provided in `.csv` format from the borealis Dryad data set interface associated with this paper is titled here as, `01_pond_data_given.` It has not been altered since download and needs to be cleaned. \

The article made frequent mention of data in a supplementary file which was a Word document that contained the following data table loaded here as, `2_suppl_data_given_totals.` It is the total biomass of each study counterpart (consumer, producer, carbon flux) against each treatment type (control, nutrient addition only, addition of a zooplanktivorous predator, 3C increase in temperature, nutrient + fish addition, nutrient + warming addition, predator + warming addition, and nutrient + predator + warming addition). At the time I wasn't sure if it was needed, but it was loaded just in case. It is known as the 'easy way' here throughout this assignment. 

```{r Loading Data, include=TRUE}
pond_data_raw=read.csv(here("01_rawdata", "1_pond_data_given.csv"))   #load in datasets
treatment_totals=read.csv(here("01_rawdata", "2_suppl_data_given_totals.csv"))  
```

________________________________________________________________________________
## 2. Manipulating Data: \

We do not know if the values were (presumably) added together first then log transformed. Likewise, we seem to know very little of what values the authors actually used for any of these analyses... means, sums, raw biomass, or log transformed- any of these combinations. Extremely frustrating. 
- *Side note: I actually did figure out what the authors ended up doing (for the most part). I was just not working the right way, instead of (obviously) working backwards, I was trying to work forwards like them, except I have to recreate things and didn't have enough information to work in this direction.*

We know that based on the Table. 1 summary statistics table that the F-value is F1,32 which should be broken down as such:
- Ambient vs. warmed are two parameters so when when one degree of freedom is subtracted it equals one (F1,x)
- Forty tanks in total then subtracting total number of tested parameters (eight) equals thirty-two so F1,32

So, I figure that it must be a statistical difference test in between the ambient and warmed if the F-value of 1,32 is only for 40 samples... we also know that these values are supposedly from the linear mixed-effects models (so we need to compute it this way first), and based on the figures the data for consumers/primary has been naturally log transformed. This is a little confusing, but another clue to try out (note: the author's reference this similar or exact study conducted in another paper by them, in which the ***biomass data was log-e transformed prior to LME regression analysis...***) 
- The author's illustrate CO2 flux as a mean with no log transformations, total consumer and total primary biomass are averaged based on every single treatment condition (eight), then naturally log transformed. I had assumed that they used these values in the LME, however, this is incorrect. While we can see that they used the data in these forms to illustrate the within their Figure 1 plots, they actually used the raw data within the LME model a.) to generate these mean + se values (this is where the easy way summary table comes from), and b.) we know that date and mesocosms were used as random factors within the LME, thus, they must be included within the original LME. 



### Easy Way:
Since I rendered and compared the values in excel and they (mostly) match this supplementary table provided, I will recreate the original plot using this data first. 
*Side note: you should come back and see if you can get them to match in R too, but I am avoiding this because it is such an absolute pain (function creation).*

```{r Generate Easy Data table 1, include=TRUE}
#Note: the values in this data table are MEAN values already computed!!!!
treatment_totals=treatment_totals %>%      #overwrite previous dataframe
  clean_names() %>%                        #clean col names
  rename(consumer_biomass=consumer_biomass_g_c_m_2, #rename them, bc long and annoying
         primary_biomass=primary_biomass_g_c_m_2, 
         co2_flux=co2_flux_mg_c_m_2_d_1) %>%
#Move se in brackets into their own col. The value "0.22" had a space behind it afterwards [unique(treatment_totals$consumer_biomass)] which is why log transformations didn't work + registering as character class types idk why
 mutate(se_consumer_biomass=str_extract(consumer_biomass, "\\(.*?\\)"),
        consumer_biomass=str_remove(consumer_biomass, "\\s*\\(.*?\\)")) %>%
 mutate(se_primary_biomass=str_extract(primary_biomass, "\\(.*?\\)"),
        primary_biomass=str_remove(primary_biomass, "\\s*\\(.*?\\)")) %>%
 mutate(se_co2=str_extract(co2_flux, "\\(.*?\\)"),
        co2_flux=str_remove(co2_flux, "\\s*\\(.*?\\)")) %>%
#Drop white space after last number
 mutate(consumer_biomass=str_trim(consumer_biomass)) %>% 
 mutate(se_co2=str_trim(se_co2)) %>% 
#Remove the brackets from the newly created standard error columns we just made:
 mutate(se_consumer_biomass=str_replace_all(se_consumer_biomass, "\\(|\\)", "")) %>%
 mutate(se_primary_biomass=str_replace_all(se_primary_biomass, "\\(|\\)", "")) %>%
 mutate(se_co2=str_replace_all(se_co2, "\\(|\\)", "")) %>%
#Make the four cols into numeric class types for when we need to compute them:
 mutate(consumer_biomass=as.numeric(consumer_biomass)) %>% 
 mutate(primary_biomass=as.numeric(primary_biomass)) %>%
 mutate(co2_flux=as.numeric(co2_flux)) %>%
 mutate(se_co2=as.numeric(se_co2)) %>%
#Log-e transform only two biomass cols:
 mutate(consumer_log=log(consumer_biomass)) %>%  #new cols for log-e, not co2 flux tho
 mutate(primary_log=log(primary_biomass)) %>%
#I need to be able to plot stuff based on if there is a predator or warming present, so we need to make something that can define that. Here, a `case_when` is easiest because there are only 4 specific circumstances. Otherwise, use an`if_else` for `str_detect`
 mutate(predator= case_when(
   treatment %in% c("P", "NP", "PW", "NPW")~"Predator Present",
   TRUE~"Predator Absent")) %>%
 mutate(temperature= case_when(
   treatment %in% c("W", "NW", "PW", "NPW")~"Warmed",
   TRUE~"Ambient")) %>%
 mutate(nutrient_case= case_when(
   treatment %in% c("N", "NP", "NW", "NPW")~"Nutrients",
   TRUE~"No Nutrients")) 
```

Personally, the long list of cleaning the `treatment_totals` data set was becoming a little long, and in order to avoid wiping my environment clean to keep running/checking changes I just started a new code chunk. 

```{r Generate Easy Data table 2, include=TRUE}
#Purpose of this chunk to to make the data frame long...  but not smartly? Just kidding, smartly. Come back and shove co2_flux if griding both plots doesn't work > did not do. 
treatment_totals_long=treatment_totals %>%
  pivot_longer(cols=c(consumer_biomass, primary_biomass), 
               names_to="biomass", 
               values_to="biomass_value") %>%
  pivot_longer(cols=c(consumer_log, primary_log), 
               names_to="biomass_log", 
               values_to="biomass_log_value") %>%
  pivot_longer(cols=c(se_consumer_biomass, se_primary_biomass), 
               names_to="standard_error", 
               values_to="standard_error_value") %>%
  mutate(standard_error_value=as.numeric(standard_error_value)) %>%
  mutate(standard_error_log=log(standard_error_value)) #checked and did correctly
```


### Hard Way:

We need to be able to use the raw data given, but it is not clearly identified by treatment type, which is a variable we will later need to run our linear regressions or plots by. 
- Issue 01: While the tank names are provided, we don't know with exact certainty which tanks were delegated as controls, and they have been randomly assigned. Here, we assume that tanks with no nutrient, fish, or warming are controls. We are basing the different treatment types listed above based on these column parameters (ex. no nutrients, no fish, but warming = 'W'). 

```{r Quick Maths 1}
#Wanted to quickly see if the number of treatments based on the the column conditions would be able to return a.) the correct number of treatment combinations, b.) if there were the right number of tanks in total (40 per sampling period x 2 = 80n total) and then c.) confirm that there should be 5 tanks per treatment type per sample period
pond_data_raw %>%
  group_by(Nutrients, Fish, Warming) %>%
  count()

######Returned in this order:#######
#control (C)
#3C increase in temperature (W) 
#addition of a zooplanktivorous predator (P)
#predator + warming addition (PW)
#nutrient addition only (N)
#nutrient + warming addition (NW)
#nutrient + fish addition (NP)
#nutrient + predator + warming addition (NPW) 
```

A function was developed in order to more quickly determine each tank treatment type based on the columns 'nutrients,' 'fish,' and 'warming' then enter in those results according to denoted treatment type. 

```{r Develop a Function to Determine Treatment Type, inclue=TRUE}
get_treatment=function(nutrients, fish, warming) 
{ #Control (C):
  if (nutrients=="no" & fish=="no" & warming=="no") { 
    return("C")              
  #Warming (W):
  } else if (nutrients=="no" & fish=="no" & warming=="yes") {
    return("W") 
  #Predator only (P): 
  } else if (nutrients=="no" & fish=="yes" & warming=="no") {
    return("P") 
  #Predator + warming (PW): 
  } else if (nutrients=="no" & fish=="yes" & warming=="yes") {
    return("P:W") 
  #Nutrient only (N): 
  } else if (nutrients=="yes" & fish=="no" & warming=="no") {
    return("N")  
  #Nutrient + warming (NW): 
  } else if (nutrients=="yes" & fish=="no" & warming=="yes") {
    return("N:W") 
  #Nutrient + Predator (NP): 
  } else if (nutrients=="yes" & fish=="yes" & warming=="no") {
    return("N:P")    
  #Nutrient + Predator + warming (NPW): 
  } else if (nutrients=="yes" & fish=="yes" & warming=="yes") {
    return("N:P:W")    
  #Check to see if any outliers and that function worked:
  } else {
    return(NA) #when ran function on df did not return any, seems to have worked
  }
}
```

```{r Generate Hard Data Table 1, include=TRUE}
pond_data_clean=pond_data_raw %>%                     #create new callable object
  clean_names() %>%                                   #edit col heading titles
  rename(consumer_biomass=total_consumer_g_c_m2,      #rename bc long/annoying
         primary_biomass=total_primary_g_c_m2, 
         co2_flux=co2_flux_mg_c_m2_d_1)  %>%
#create new columns from function based on treatment type:
 rowwise() %>%
 mutate(treatment=get_treatment(nutrients, fish, warming)) %>% 
 ungroup()
#Visually checked but also ran a quick count by treatment type and returned the same values (n=10; n=5 per sample period) for 8 treatment types. 
```

```{r Generate Hard Data Table 2, include=TRUE}
pond_data_clean_long=pond_data_clean %>%    #create new object
  mutate(date=as.factor(date)) %>%          #random factors need to be factors
  mutate(tank=as.factor(tank)) %>%          #random factors need to be factors
  mutate(treatment=as.factor(treatment)) %>%#categorical needs to be a factor
  group_by(treatment) %>%                   #group each mean values by 'treatment' type
#Make raw biomass values into means per treatment type (total table values match)
  mutate(consumer_bio_mean=round(sum(consumer_biomass)/10, 2)) %>% #10 samples per trt
  mutate(primary_bio_mean=round(sum(primary_biomass)/10, 2)) %>%
  mutate(co2_flux_mean=round(sum(co2_flux)/10, 2)) %>%
  ungroup() %>%
#Make raw biomass data into log transformed values ONLY, NOT carbon since we will not need it for the LME and it conveniently produces an error here *delete*
  mutate(consumer_bio_log=round(log(consumer_biomass),2)) %>%
  mutate(primary_bio_log=round(log(primary_biomass),2)) %>%
#pivot this table longer so as to better filter for all the different LMEs:
  pivot_longer(cols=c(consumer_biomass, primary_biomass, co2_flux), 
               names_to="category", 
               values_to="category_value") %>%
  pivot_longer(cols=c(consumer_bio_mean, primary_bio_mean, co2_flux_mean), 
               names_to="category_mean", 
               values_to="category_mean_value") %>%
  pivot_longer(cols=c(consumer_bio_log, primary_bio_log), 
               names_to="biomass_log", 
               values_to="biomass_log_values")
```

________________________________________________________________________________
## 3. Statistical Analyses on Data: \

### Easy Way: \
There isn't really an easy way to confirm the SE values provided from the total_treatment` summary table, since we require all of the consumer, primary, and CO2 raw data in order to compute these parameters. We also cannot run any modeling on this table since, again, we require the raw data values from all mesocosms. 


### Hard Way: \
Here, this is the raw data (`pond_data_given`) that the authors used to render the mean and SE values within the `total_treatments` summary table. We must run the model and receive these similar values using this original data table. 
To be frank, I was terrified of generating a function for a process which I was already hesitant and shakey to begin with, so I unfortunately brute forced all of these LME things. 


allowing for nested random effects. The within-group errors are allowed to be correlated and/or have unequal variances.
- did they compare with the control? thats where the F stat comes from? two groups (treatment N vs control) for consumers? 
- i first tried running with just the raw biomass values, but previous paper said that they first log transformed biomass data prior to running the LME... fuck ugh
- tbh i dont get how we can see if the fixed variable impacts the response because the fixed is a treatment, not changing over time (not even a number)
- tank and date are in fact treated as if they are crossed because they are: two of the same tanks appear on the two same dates (a given factor appears in more than one level of another factor)
- upper ci and lower ci are different: what did they actually do? means? then apply to both? wher are these ci intervals, and what test__> lme predict
=anova consumer, primary, co2 somehow? 

lmer(testScore ~ bodyLength2 + (1|mountainRange) + (1|site), data = dragons) 
tank 
```{r Developing Linear Models: Consumers, include=TRUE}
#All LME models (treatment types) on consumer biomass (C,N,P,W,N:P,W:N,W:P,W:P:N) currently on raw biomass data.

#MODELING: N on raw consumer biomass ------------------------------------------
model_CB_N=pond_data_clean_long %>%               #df for consumer data, N only
  select(-year, -nutrients, -fish, -warming) %>%
  filter(category=="consumer_biomass") %>%
  filter(category_mean=="consumer_bio_mean") %>%
  filter(biomass_log=="consumer_bio_log") %>%
  filter(treatment=="N")
#fixed=treatment (N) // response=consumers // randoms=date+tank
#order: what we wanna know (response)~what should impact this (depend) + randoms
#here `category_value` can be entered because we specifically filtered it above
#model_CBraw_N=lmer(category_value~treatment+(1|tank)+(1|date), data=model_CB_N)


model_CB=pond_data_clean_long %>%               #df for consumer data, N only
  select(-year, -nutrients, -fish, -warming) %>%
  filter(category=="consumer_biomass") %>%
  filter(category_mean=="consumer_bio_mean") %>%
  filter(biomass_log=="consumer_bio_log")

model_CBraw=lmer(category_value~treatment+(1|tank)+(1|date), data=model_CB)
model_CBlog=lmer(biomass_log_values~treatment+(1|tank)+(1|date), data=model_CB)
```

```{r}
emmeans_model_CB=emmeans(model_CBraw, specs = pairwise ~ treatment)
summary(emmeans_model_CB)
```
```{r}
model_CB_lme=as.lme(model_CBraw)
se_predictions=predictSE.lme(object = model_CB_lme)
print(se_predictions)
```


```{r}
model_CBraw_summary=summary(model_CBraw)
summary(model_CBraw)
#summary(model_CBlog)
anova(model_CBraw)
```



```{r}
model_CB %>%
  print()
```


lmerTest

                     




________________________________________________________________________________
## 4. Plotting Data: \

### Easy Way: \
I first attempted to plot the biomass data straight into a `facet_grid` which worked, but the all the different plot axes seemed very close together, and I was not able to make them as large as I had wished. The thought here was to plot logged biomass data and carbon flux data, then plop the former on-top of the latter later down the road.
However, I then just plotted each figure separately based on filtering conditions and merged them all together. \

```{r Easy Way Plot 1, include=FALSE}
easy_biomass_plots=treatment_totals_long %>%
#Generate the plot:  
  ggplot() +
  facet_grid(biomass_log~predator, space="free") +
  geom_point(aes(x=temperature, y=biomass_log_value, shape=factor(nutrient_case))) +
#Plot aesthetics:
  theme_minimal() +  
  theme(legend.position="bottom")
  labs(y="New Y-Axis Title") 
print(easy_biomass_plots)
```


- Keep grid lines because even though lots of small confusing plots that are harder to read, they are small, so adding so many lines makes things more cluttered?
- error bars and maybe points seem to not quite be lining up with given plots, checking for that error. Ok so, I think that the authors are making the points too large which is making things look slightly off (log values are so small), since I am literally using the exact values that they GAVE as mean then logged values. For the most part these values match the ones I computed in excel (constant rounding not off by more than 0.05). Double checked that R computed log right (again) and that values match. They do, based on logging the means provided by authors.   
- same with se bars. using the values they gave for means and se bars for those means. logging the se bars... doesn't make any sense and also gives us WILD error bars (much larger than those used by the authors). wish i could say i was intelligent enough that this was obvious and easy. so, mean se bar logged because scale of data for consistency SHOULD BE USED HERE! but i did NOT to match paper aesthetics!! changing the y axis to 05->0 like in the paper actually cuts off these error bars. so they had to stay smaller.
- ok so recomputing the se values with a 95% ci based on the mean values (should be fine because we need mean anyways) 
- did these fuckers just take raw mean values and throw it on a log scale?????? no.. based on what i read i dont think you can directly chage a scale if or when data will become negative. so they mustve converted it earlier(?) they gave the data but they are making me WORK for it christ 
kay wait- thought/realization: the mean +/- se given to us in the summary table is for BOTH warm and ambient treatments???????? but if were plotting them seperately then we need the respective data for each treatment type BUT also for everything we keep breaking things down for. 


```{r Generating Figures: Easy Way Plot A, include=FALSE}
#Plot a: i.) consumer LOG biomass ii.) no predators
easy_plota_con_nopreds=treatment_totals_long %>% #checked this df combo just to be safe 
  filter(biomass=="consumer_biomass") %>%    
  filter(biomass_log=="consumer_log") %>%
  filter(standard_error=="se_consumer_biomass") %>%
  filter(predator=="Predator Absent") %>%
#Generate the plot:  
  ggplot() +
  geom_point(aes(x=temperature, y=biomass_log_value, shape=factor(nutrient_case)), 
             size=3) +                        #plot points based on strict filters above
  geom_errorbar(aes(x=temperature, y=biomass_log_value,  #SE based on grouped conditions
                ymin=biomass_log_value-standard_error_value, 
                ymax=biomass_log_value+standard_error_value,
                group=interaction(treatment, nutrient_case)), width=0.03, guide=NULL) +
  geom_line(aes(x=temperature, y=biomass_log_value, group=factor(nutrient_case), 
                linetype=factor(nutrient_case)), colour="black", guide=NULL) +
  #ylim(-5,0) +                              
  scale_y_continuous(breaks=seq(-5, 0, by=1), #define y-axis line segments
                   limits=c(-5, 0)) +         #define y-axis and                     
  scale_shape_manual(name="",                 #remove line legend title
                    values=c(19, 1),          #only way to keep lines w/o legend
                    labels=c("No Nutrients", "Nutrients")) +
#Plot aesthetics:
  theme_minimal() +                           #theme_bw() couldn't alter outer grid
  theme(plot.background=element_rect(fill="white"),   #make figure background white
        panel.grid.major=element_line(colour="white"),#make figure grid lines white
        axis.line.x=element_line(colour="black"),     #set outer x-axis line colour  
        axis.line.y=element_line(colour="black")) +   #set outer y-axis line colour 
#  theme(plot.margin = margin(0.75, 2, 0, 0.5, "cm")) +#T, R, B, L (plot margins)
  theme(legend.position="bottom") +                   #set legend to the bottom
  guides(linetype=FALSE) +                            #get rid of line legend omfg
#Generate labels:    
  labs(x=NULL,                                        #no x-axis label here
       y=expression(atop(log[e]~"consumer biomass"~(g~C~m^{-2})))) +
        #formulate y-axis label expression, add the common back where you want the break
  theme(axis.title.y=element_text(vjust=-5)) +       #move y-axis title closer right
  labs(title="Predator Absent") +                    #only add plot title 1/2
  theme(plot.title=element_text(hjust=.50))          

print(easy_plota_con_nopreds)
```

```{r Generating Figures: Easy Way Plot B, include=FALSE}
#Plot b: i.) consumer LOG biomass ii.) predators 
easy_plotb_con_preds=treatment_totals_long %>% #checked this df combo just to be safe 
  filter(biomass=="consumer_biomass") %>%    
  filter(biomass_log=="consumer_log") %>%
  filter(standard_error=="se_consumer_biomass") %>%
  filter(predator=="Predator Present") %>%                             #changed 
#Generate the plot:  
  ggplot() +
  geom_point(aes(x=temperature, y=biomass_log_value, shape=factor(nutrient_case)), 
             size=3) +                        #plot points based on strict filters above
  geom_errorbar(aes(x=temperature, y=biomass_log_value,  #SE based on grouped conditions
                ymin=biomass_log_value-standard_error_value, 
                ymax=biomass_log_value+standard_error_value,
                group=interaction(treatment, nutrient_case)), width=0.03, guide=NULL) +
  geom_line(aes(x=temperature, y=biomass_log_value, group=factor(nutrient_case), 
                linetype=factor(nutrient_case)), colour="black", guide=NULL) +
  #ylim(-5,0) +                              
  scale_y_continuous(breaks=seq(-5, 0, by=1), #define y-axis line segments
                   limits=c(-5, 0)) +         #define y-axis and                     
  scale_shape_manual(name="",                 #remove line legend title
                    values=c(19, 1),          #only way to keep lines w/o legend
                    labels=c("No Nutrients", "Nutrients")) +
#Plot aesthetics:
  theme_minimal() +                           #theme_bw() couldn't alter outer grid
  theme(plot.background=element_rect(fill="white"),   #make figure background white
        panel.grid.major=element_line(colour="white"),#make figure grid lines white
        axis.line.x=element_line(colour="black"),     #set outer x-axis line colour  
        axis.line.y=element_line(colour="black")) +   #set outer y-axis line colour 
#  theme(plot.margin = margin(0.75, 2, 0, 0.5, "cm")) +#T, R, B, L (plot margins)
  theme(legend.position="bottom") +                   #set legend to the bottom
  guides(linetype=FALSE) +                            #get rid of line legend omfg
#Generate labels:    
  labs(x=NULL, y=NULL) +                              #no x-axis label here
#     y=expression(atop(log[e]~"consumer",            #formulate y-axis label expression
#                         "biomass"~(g~C~m^{-2}))))
  labs(title="Predator Present") +                    #only add plot title 2/2
  theme(plot.title=element_text(hjust=.50))          

print(easy_plotb_con_preds)
```

```{r Generating Figures: Easy Way Plot C, include=FALSE}
#Plot c: i.) primary LOG biomass ii.) no predators 
easy_plotc_prim_nopreds=treatment_totals_long %>%  
  filter(biomass=="primary_biomass") %>%    
  filter(biomass_log=="primary_log") %>%
  filter(standard_error=="se_primary_biomass") %>%
  filter(predator=="Predator Absent") %>%                             
#Generate the plot:  
  ggplot() +
  geom_point(aes(x=temperature, y=biomass_log_value, shape=factor(nutrient_case)), 
             size=3) +                        #plot points based on strict filters above
  geom_errorbar(aes(x=temperature, y=biomass_log_value,  #SE based on grouped conditions
                ymin=biomass_log_value-standard_error_value, 
                ymax=biomass_log_value+standard_error_value,
                group=interaction(treatment, nutrient_case)), width=0.03, guide=NULL) +
  geom_line(aes(x=temperature, y=biomass_log_value, group=factor(nutrient_case), 
                linetype=factor(nutrient_case)), colour="black", guide=NULL) +
  #ylim(-5,0) +                              
  scale_y_continuous(breaks=seq(-2, 3, by=1), #define y-axis line segments
                   limits=c(-2, 3)) +         #define y-axis and                     
  scale_shape_manual(name="",                 #remove line legend title
                    values=c(19, 1),          #only way to keep lines w/o legend
                    labels=c("No Nutrients", "Nutrients")) +
#Plot aesthetics:
  theme_minimal() +                           #theme_bw() couldn't alter outer grid
  theme(plot.background=element_rect(fill="white"),   #make figure background white
        panel.grid.major=element_line(colour="white"),#make figure grid lines white
        axis.line.x=element_line(colour="black"),     #set outer x-axis line colour  
        axis.line.y=element_line(colour="black")) +   #set outer y-axis line colour 
#  theme(plot.margin = margin(0.75, 2, 0, 0.5, "cm")) +#T, R, B, L (plot margins)
  theme(legend.position="bottom") +                   #set legend to the bottom
  guides(linetype=FALSE) +                            #get rid of line legend omfg
#Generate labels:    
  labs(x=NULL,                                        #no x-axis label here
       y=expression(atop(log[e]~"primary producer biomass"~(g~C~m^{-2})))) +
     #formulate y-axis label expression, add the common back where you want the break
  theme(axis.title.y=element_text(vjust =-5))         #move y-axis title closer right
print(easy_plotc_prim_nopreds)
```
```{r Generating Figures: Easy Way Plot D, include=FALSE}
#Plot d: i.) primary LOG biomass ii.) predators 
easy_plotd_prim_preds=treatment_totals_long %>%  
  filter(biomass=="primary_biomass") %>%    
  filter(biomass_log=="primary_log") %>%
  filter(standard_error=="se_primary_biomass") %>%
  filter(predator=="Predator Present") %>%                             
#Generate the plot:  
  ggplot() +
  geom_point(aes(x=temperature, y=biomass_log_value, shape=factor(nutrient_case)), 
             size=3) +                        #plot points based on strict filters above
  geom_errorbar(aes(x=temperature, y=biomass_log_value,  #SE based on grouped conditions
                ymin=biomass_log_value-standard_error_value, 
                ymax=biomass_log_value+standard_error_value,
                group=interaction(treatment, nutrient_case)), width=0.03, guide=NULL) +
  geom_line(aes(x=temperature, y=biomass_log_value, group=factor(nutrient_case), 
                linetype=factor(nutrient_case)), colour="black", guide=NULL) +
  #ylim(-5,0) +                              
  scale_y_continuous(breaks=seq(-2, 3, by=1), #define y-axis line segments
                   limits=c(-2, 3)) +         #define y-axis and                     
  scale_shape_manual(name="",                 #remove line legend title
                    values=c(19, 1),          #only way to keep lines w/o legend
                    labels=c("No Nutrients", "Nutrients")) +
#Plot aesthetics:
  theme_minimal() +                           #theme_bw() couldn't alter outer grid
  theme(plot.background=element_rect(fill="white"),   #make figure background white
        panel.grid.major=element_line(colour="white"),#make figure grid lines white
        axis.line.x=element_line(colour="black"),     #set outer x-axis line colour  
        axis.line.y=element_line(colour="black")) +   #set outer y-axis line colour 
#  theme(plot.margin = margin(0.75, 2, 0, 0.5, "cm")) +#T, R, B, L (plot margins)
  theme(legend.position="bottom") +                   #set legend to the bottom
  guides(linetype=FALSE) +                            #get rid of line legend omfg
#Generate labels:    
  labs(x=NULL, y=NULL)                                #no x-axis label here
#      y=expression(atop(log[e]~"primary producer biomass"~(g~C~m^{-2}))))
       #formulate y-axis label expression, add the common back where you want the break

print(easy_plotd_prim_preds)
#The standard SE value for ambient nutrient is 3.21 which is why it wont plot here..
#there is something defintely wrong with SE values here and i dont know what
```
```{r Generating Figures: Easy Way Plot E, include=FALSE}
#Plot e: i.) just c02 flux mean values ii.) no predators 
easy_plote_co2_nopreds=treatment_totals_long %>%  
  filter(predator=="Predator Absent") %>%                             
#Generate the plot:  
  ggplot() +
  geom_point(aes(x=temperature, y=co2_flux, shape=factor(nutrient_case)), size=3) +     
  geom_errorbar(aes(x=temperature, y=co2_flux,  #SE based on grouped conditions
                ymin=co2_flux-se_co2, ymax=co2_flux+se_co2,
                group=interaction(treatment, nutrient_case)), width=0.03, guide=NULL) +
  geom_line(aes(x=temperature, y=co2_flux, group=factor(nutrient_case), 
                linetype=factor(nutrient_case)), colour="black", guide=NULL) +
  scale_y_continuous(breaks=seq(-100, 50, by=50), #define y-axis line segments
                   limits=c(-110, 50)) +          #define y-axis                   
  scale_shape_manual(name="",                     #remove line legend title
                    values=c(19, 1),              #only way to keep lines w/o legend
                    labels=c("No Nutrients", "Nutrients")) +
  geom_hline(yintercept=0, linetype="longdash") + #add co2 influx/efflux line
#Plot aesthetics:
  theme_minimal() +                           #theme_bw() couldn't alter outer grid
  theme(plot.background=element_rect(fill="white"),   #make figure background white
        panel.grid.major=element_line(colour="white"),#make figure grid lines white
        axis.line.x=element_line(colour="black"),     #set outer x-axis line colour  
        axis.line.y=element_line(colour="black")) +   #set outer y-axis line colour 
#  theme(plot.margin = margin(0.75, 2, 0, 0.5, "cm")) +#T, R, B, L (plot margins)
  theme(legend.position="bottom") +                   #set legend to the bottom
  guides(linetype=FALSE) +                            #get rid of line legend omfg
#Generate labels:    
  labs(x=NULL,                                        #no x-axis label here
       y=expression(atop(CO[2]~flux~(mg~C~m^{-2}~d^{-1})))) +
       #formulate y-axis label expression
 theme(axis.title.y=element_text(vjust =-7))         #move y-axis title closer right

print(easy_plote_co2_nopreds)
```
```{r Generating Figures: Easy Way Plot F, include=FALSE}
#Plot f: i.) just c02 flux mean values ii.) predators 
easy_plotf_co2_preds=treatment_totals_long %>%  
  filter(predator=="Predator Present") %>%                             
#Generate the plot:  
  ggplot() +
  geom_point(aes(x=temperature, y=co2_flux, shape=factor(nutrient_case)), size=3) +     
  geom_errorbar(aes(x=temperature, y=co2_flux,  #SE based on grouped conditions
                ymin=co2_flux-se_co2, ymax=co2_flux+se_co2,
                group=interaction(treatment, nutrient_case)), width=0.03, guide=NULL) +
  geom_line(aes(x=temperature, y=co2_flux, group=factor(nutrient_case), 
                linetype=factor(nutrient_case)), colour="black", guide=NULL) +
  scale_y_continuous(breaks=seq(-100, 50, by=50), #define y-axis line segments
                   limits=c(-110, 50)) +          #define y-axis                   
  scale_shape_manual(name="",                     #remove line legend title
                    values=c(19, 1),              #only way to keep lines w/o legend
                    labels=c("No Nutrients", "Nutrients")) +
  geom_hline(yintercept=0, linetype="longdash") + #add co2 influx/efflux line
#Plot aesthetics:
  theme_minimal() +                           #theme_bw() couldn't alter outer grid
  theme(plot.background=element_rect(fill="white"),   #make figure background white
        panel.grid.major=element_line(colour="white"),#make figure grid lines white
        axis.line.x=element_line(colour="black"),     #set outer x-axis line colour  
        axis.line.y=element_line(colour="black")) +   #set outer y-axis line colour 
#  theme(plot.margin = margin(0.75, 2, 0, 0.5, "cm")) +#T, R, B, L (plot margins)
  theme(legend.position="bottom") +                   #set legend to the bottom
  guides(linetype=FALSE) +                            #get rid of line legend omfg
#Generate labels:    
  labs(x=NULL, y=NULL)                                        #no x-axis label here
#       y=expression(atop(CO[2]~flux,(mg~C~m^{-2}~d^{-1}))))
    #formulate y-axis label expression, add the common back where you want the break
print(easy_plotf_co2_preds)
```
 
So. Since I am a *fraud* and could not for the life of me perfect the merging of these plots in a more humane and logical manner, I had to cheat and merge them in groups of two to preserve the overall plot readability size. I also could not get the outter black border to cease to exist, due to no fault of `patchwork()` or `cowplot()` or `ggpubr` from a package that I forget, but I don't know. The fact that different functions can alter these plots, but by altering the margins is the only way to remove them means it *must* be something with how they are getting lined up. But after an embarrassing amount of time, it is sadly time to move on. 

```{r Merge Easy Plots AB, include}
easy_plot_ab=                                         #create new object
  easy_plota_con_nopreds+easy_plotb_con_preds +       #here adding, but '|' could work
    plot_layout(guides="collect") +                   #merge legends
    plot_annotation(tag_levels=c('a','b')) &          #symbol applies to all plots
  theme(
    legend.position="top",                            #move legend to the top
    legend.direction="vertical",                      #make points show up and down 
    legend.justification="right",                     #move legend to the right
    panel.border=element_blank())                     #remove extra grid lines

print(easy_plot_ab)
```
```{r Merge Easy Plots CD, include}
easy_plot_cd=                                         #create new object
  easy_plotc_prim_nopreds+easy_plotd_prim_preds +     #here adding, but '|' could work
    plot_layout(guides="collect") +                   #merge legends
    plot_annotation(tag_levels=list(c('c','d'))) &    #symbol applies to all plots
  theme(legend.position="none",                       #remove legends from plot
    panel.border=element_blank(),                     #remove extra grid lines
    plot.margin=margin(1.75, 0.20, 0.20, 0.20, "cm"))#make size match first row
  
print(easy_plot_cd)
```
```{r Merge Easy Plots EF, include}
easy_plot_ef=                                         #create new object
  easy_plote_co2_nopreds+easy_plotf_co2_preds +       #here adding, but '|' could work
    plot_layout(guides="collect") +                   #merge legends
    plot_annotation(tag_levels=list(c('e','f'))) &    #symbol applies to all plots
  theme(legend.position="none",                       #remove legends from plot
    panel.border=element_blank(),                     #remove extra grid lines
    plot.margin=margin(1.75, 0.20, 0.20, 0.20, "cm"))#make size match first row
  
print(easy_plot_ef)
```





